N-dimensional Tic Tac Toe, and Adventure in Modules

### 2017-05-12 paper/paper.tex
structural typing section

	\section{Design Decisions}
  % (Most important!) Describe in detail 3-5 design decisions you made during your
  % project, with respect to abstraction and/or modularity. For example, what is a
  % piece of information that you chose to hide and why? How did you decide on a
  % particular way of separating your system into modules compared to another. How
  % did you reduce coupling between two modules? These design decisions should
  % emphasize how you applied ideas from the papers we’ve read (and associated
  % discussions) to the design of your project. It might be helpful here to
  % include snippets from old versions (or hypothetical alternative versions) of
  % your project to compare and contrast with the final version.

  \subsection{Creating an Abstract Game Engine}
  % How we came up with the abstract design
  The design for our game engine stemmed from trying to isolate the common
  general components of what would be required to create a running game.
  Rather than focus on creating one specific game implementation - we wanted to
  look at the larger picture and view the problem in terms of the common pattern
  that emerged between these kinds of games.

  Each of the portions of the \textit{State}-\textit{Actions}-\textit{Agents}
  design is independent so that - once created - it does not rely on other
  modules for it's internal pieces.
  Each of the constituent parts of the game engine is more akin to a node on a
  graph and the interfaces are like edges.
  The pieces should be thought of as islands of code that could then interact
  with other modules. By using the game model to define a specific game - we
  essentially create more specific versions of those islands following the
  recipe spelled out in the signatures.

  \subsection{Higher Ordered Signatures, and the ``Include'' incantation}
  % Something about how previously we just had signature -> structure -> functors
  % but with include it became possible to 
  \subsection{Separation of IO, or How I learned to not fight SML in search of Purity}
  In the first implementation of our project we simulated Haskell's IO monad by
  creating a ``show'' signature, and an ``IO'' functor, as shown below:

>>>>

	\section{Design Decisions}
  % (Most important!) Describe in detail 3-5 design decisions you made during your
  % project, with respect to abstraction and/or modularity. For example, what is a
  % piece of information that you chose to hide and why? How did you decide on a
  % particular way of separating your system into modules compared to another. How
  % did you reduce coupling between two modules? These design decisions should
  % emphasize how you applied ideas from the papers we’ve read (and associated
  % discussions) to the design of your project. It might be helpful here to
  % include snippets from old versions (or hypothetical alternative versions) of
  % your project to compare and contrast with the final version.
  \subsection{Creating an Abstract Game Engine}
  % How we came up with the abstract design
  \subsection{Exploiting Structural Typing}
  % Something about how previously we just had signature -> structure -> functors
  % but with include it became possible to
  One major element of SML's module system that we were about to exploit in our project was
  the structural matching of modules to signatures. Coupled with liberal use of the \texttt{include}
  keyword, this allowed us to easily extend signature definitions in a way reminiscent of inheritance
  and subclassing in OOP.

  This design decision is most apparent in our extension of the abstract game engine to
  our chosen implementation domain of \textit{Tic-tac-toe}. Our more abstract interfaces defined
  in \texttt{game.sml} were often insufficient for describing the full interface
  needed for our concrete implementation. For example, \textit{Tic-tac-toe} state modules
  needed to export a \texttt{cell} datatype that represented the actual Xs and Os placed on the game
  board. Including this datatype in the abstract \texttt{STATE} signature in \texttt{game.sml}
  would be inappropriate, as any game outside of \textit{Tic-tac-toe} has no need for the
  \texttt{cell} datatype. Instead, we created a new signature \texttt{TTTSTATE} that extended
  the \texttt{STATE} signature via the \texttt{include} keyword with additional interface elements.
  Additionally, this extension refined
  some of the abstract types in \texttt{STATE} using the \texttt{where} keyword in order to give them
  concrete type definitions. For example, the abstract type \texttt{state} is defined to always be
  equivalent to the type \texttt{cell * (cell Matrix.container)} in a module implementing \texttt{TTTSTATE}.

  We were able to exploit this refined state signature to narrow the domain of the \texttt{TttActionFn}
  functor to only states appropriate for \textit{Tic-tac-toe}, rather than any abstract game state.
  However, thanks to the structural matching properties of the SML module system, we can still
  use a module implementing \texttt{TTTSTATE} wherever we would use a module implementing
  \texttt{STATE}. In fact, because the matching is structural and not nominal, we get this property
  for free, with no need to specify the relationship by name in the code. This allows us to still use
  signatures and functors in our implementation that depend only on the more abstract 
  definitions like the \texttt{AGENT} signature
  or the \texttt{ExecFn} functor, which are only defined for the more abstract game engine.

  \subsection{Separation of IO, or How I learned to not fight SML in search of Purity}
  In the first implementation of our project we simulated Haskell's IO monad by
  creating a ``show'' signature, and an ``IO'' functor, as shown below:
